---
title: "pml_assignment"
author: "Grissel Priyanka Mathias"
date: "10/18/2020"
output: html_document
---
#Introduction

To collect all out physical activity with ease we use fitness band. We collect data from accelometers of total of 6 people. They do the lifts in 5 correct and incorrect way. This is a part of week 4 practical machine learning peer graded assignment.

#libraries

```{r warning=FALSE, error=FALSE}
library(rattle)
library(caret)
library(rpart)
library(rpart.plot)
library(corrplot)
library(randomForest)
library(RColorBrewer)
library(xgboost)
library(dplyr)
```

# Data Extraction

```{r}
d_directory = "./data"
trainingURL = "https://d396qusza40orc.cloudfront.net/predmachlearn/pml-training.csv"
testingURL = "https://d396qusza40orc.cloudfront.net/predmachlearn/pml-testing.csv"
TrainFile = "pml-training.csv"
TestFile = "pml-Testing.csv"
if (!file.exists(d_directory)) {
  dir.create(d_directory)
}
if (!file.exists(file.path(d_directory, TrainFile))) {
  download.file(trainingURL, destfile=file.path(d_directory, TrainFile))
}
if (!file.exists(file.path(d_directory, TestFile))) {
  download.file(testingURL, destfile=file.path(d_directory, TestFile))
}
```

# Reading Data

```{r}
train <- read.csv(file.path(d_directory, TrainFile))
Testing <- read.csv(file.path(d_directory, TestFile))
dim(train)
dim(Testing)
head(train)
```

# Cleaning Data


```{r}
sum(complete.cases(train))
```

# Eliminate NA columns

```{r}
colnames(train)
plot(colMeans(is.na(train)))
```

```{r}
TrainingClass = train$classe
Train_Raw = train[, sapply(train, is.numeric)]
Test_Raw = Testing[, sapply(Testing, is.numeric)]
```

Removing columns with NA values

```{r}
TrainingFilers <- Train_Raw[, colSums(is.na(Train_Raw)) == 0]
TrainingFilers$classe = TrainingClass
TestingFilers <- Test_Raw[, colSums(is.na(Test_Raw)) == 0]
```

```{r}
dim(TrainingFilers)
dim(TestingFilers)
```

```{r}
trash = !grepl("X|timestamp", colnames(TrainingFilers))
Col = colnames(TrainingFilers)[trash]
TrainingFilers = TrainingFilers %>%
  select(Col)
trash = !grepl("X|timestamp", colnames(TestingFilers))
Col = colnames(TestingFilers)[trash]
TestingFilers = TestingFilers %>%
  select(Col)
```

```{r}
dim(TrainingFilers)
dim(TestingFilers)
```

```{r}
set.seed(12022018) # Today's date
inTraining <- createDataPartition(TrainingFilers$classe, p=0.70, list=F)
Training_Data <- TrainingFilers[inTraining, ]
Val_Data <- TrainingFilers[-inTraining, ]
dim(Training_Data)
```

#RandomForest Data Model

```{r}
ConRF <- trainControl(method="cv", 5, allowParallel = TRUE)
RFModel <- train(classe ~ ., data=Training_Data, method="rf", trControl=ConRF, ntree=250)
RFModel
```

```{r}
RF_Pre <- predict(RFModel, Val_Data)
confusionMatrix(Val_Data$classe, RF_Pre)
```

```{r}
xgb_Control <- trainControl(method="cv", 5, allowParallel = TRUE)
xgb_Model <- train(classe ~ ., data=Training_Data, method="xgbTree", trControl=xgb_Control)
```

```{r}
xgb_Model
```

#Validation performance data set

```{r}
xgb_Pre <- predict(xgb_Model, Val_Data)
confusionMatrix(Val_Data$classe, xgb_Pre)
```

# Comparing

```{r}

ModelR <- resamples(list(RF=RFModel, XGB=xgb_Model))
summary(ModelR)
bwplot(ModelR)
dotplot(ModelR)
```

```{r}
RFRes <- predict(RFModel, TestingFilers[, -length(names(TestingFilers))])
xgbRes <- predict(xgb_Model, TestingFilers[, -length(names(TestingFilers))])
RFRes
xgbRes
confusionMatrix(RFRes, xgbRes)
```